---
title: "R introduction "
output: 
  learnr::tutorial:
    theme: "journal"
    css: "css/mystyle.css"
runtime: shiny_prerendered
---
  
```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(shinythemes)
knitr::opts_chunk$set(echo = FALSE)
```





## Welcome!

Hi! This is an introduction for R written for  Economic Data students at the University of Bristol. (But anyone is welcome to use it.) The goal of the tutorial is to create data visualizations of income inequality using data from the [World Inequality Database](https://wid.world/). To achieve this goal we'll go through eight small "lectures" and cover the following topics:


* plotting data
* loading data
* adjusting charts
* merging data
* modifying data
* merging data
* appending data
* restructuring data






The tutorial includes small code blocks. The code blocks allow you to immediately try the described functions and observe the result. There are also small quizzes to check your progress and assess whether the code ran correctly.

Let's get started. Clcik on  <tt>Next topic</tt> to start the first lecture.

Best,
Hans

by  Hans H. Sievertsen, University of Bristol 13/2/2020





## 1 Create a chart 

*In this section we load data into R and create a chart using this data.*
  
  Let's say that we would like to create a line chart of the income shares in the United Kingdom. In Microsoft Excel we would do this using the following steps:

1. Open the data in Microsoft Excel
2. Select "Insert"
3. Select "Line chart"
4. Select "Select Data"
5. Select the data for the horizontal axis 
6. Select the data for the vertical axis 

In R we can load data and create a line chart by asking R to run two lines of "code". A code is basically just written instructions that replace our "mouse clicking" in Microsoft Excel. So instead of using the mouse to click on "File" and then "Open" and then "Browse" and  so on, we use the "R language" to tell R to do this.  The box below contains your first R code. In programming-based software we first write some code and then tell the software to "execute" or "run" this code.  You can run the code below  by clicking "Run Code". Try it!

```{r first-chart1, exercise=TRUE,  warning=FALSE, message=FALSE,cache=TRUE }
wid_data<-read_csv("https://www.hhsievertsen.net/economicdata/src/wid_uk_t1.csv")
ggplot(data=wid_data,aes(x=Year,y=Income_share))+geom_line()

```
